---
title: "Prediction intervals for Loess forecasting (simulation-based)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Prediction intervals for Loess forecasting (simulation-based)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`ahead` is a package for **univariate** and **multivariate time series forecasting**, with uncertainty quantification (R and [Python](https://techtonique.github.io/ahead_python/)).

The model used in this demo is `stats::loess` (Local Polynomial Regression Fitting), 
**adapted to univariate forecasting** in `ahead::loessf`. 

Currently for this model (as of 2023-08-28), for **uncertainty quantification**, options are:

-   Independent **bootstrap** of the residuals
-   Multivariate circular **block bootstrap** of the residuals
-   Multivariate moving block bootstrap of the residuals
-   Adjustment of a copula to the residuals
-   **More options to come** in the future.

Please remember that _in real life_, this model's hyperparameters will have to be tuned.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Install `ahead`

Here's how to install the R version of the package:

-   **1st method**: from [R-universe](https://techtonique.r-universe.dev)

    In R console:

    ``` r
    options(repos = c(
        techtonique = 'https://techtonique.r-universe.dev',
        CRAN = 'https://cloud.r-project.org'))

    install.packages("ahead")
    ```

-   **2nd method**: from [Github](https://github.com/Techtonique/ahead)

    In R console:

    ``` r
    devtools::install_github("Techtonique/ahead")
    ```

    Or

    ``` r
    remotes::install_github("Techtonique/ahead")
    ```

And here are the packages that will be used in this vignette:

```{r include=FALSE}
library(ahead)
library(fpp)
library(datasets)
library(randomForest)
library(e1071)
```

``` r
library(ahead)
library(datasets)
```

# `ahead::loessf` on Nile dataset 

```{r "bootstrap-mean", fig.width=7.2}
plot(loessf(Nile, h=20, 
            type_pi = "bootstrap",
            type_aggregation = "mean",
            level=95, B=10))
```

```{r "bootstrap-median", fig.width=7.2}
plot(loessf(Nile, h=20, 
            type_pi = "bootstrap",
            type_aggregation = "median",
            level=95, B=10))
```

```{r "blockbootstrap-mean", fig.width=7.2}
plot(loessf(Nile, h=20, 
            type_pi = "blockbootstrap",
            type_aggregation = "mean",
            level=95, B=10))
```

```{r "blockbootstrap-median", fig.width=7.2}
plot(loessf(Nile, h=20, 
            type_pi = "blockbootstrap",
            type_aggregation = "median",
            level=95, B=10))
```

