---
title: "Risk-neutralize simulations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Risk-neutralize simulations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Contents:**

 - 0 - Install `ahead`

 - 1 - Get and transform data

 - 2 - Risk-neutralize simulations
 
 - 3 - Visualization

# 0 - Install `ahead`

`ahead` is released under the BSD Clear license. Here's how to install the R version of the package:

-   **1st method**: from [R-universe](https://techtonique.r-universe.dev)

    In R console:

    ``` r
    options(repos = c(
        techtonique = 'https://techtonique.r-universe.dev',
        CRAN = 'https://cloud.r-project.org'))

    install.packages("ahead")
    ```

-   **2nd method**: from [Github](https://github.com/Techtonique/ahead)

    In R console:

    ``` r
    devtools::install_github("Techtonique/ahead")
    ```

    Or

    ``` r
    remotes::install_github("Techtonique/ahead")
    ```

Using `ahead`:

```{r include=FALSE}
library(ahead)
```

```R
library(ahead)
library(fpp)
```

# 1 - Get and transform data

```{r}
data(EuStockMarkets)

EuStocks <- ts(EuStockMarkets[1:100, ], 
               start = start(EuStockMarkets),
               frequency = frequency(EuStockMarkets))

EuStocksLogReturns <- ahead::getreturns(EuStocks, type = "log")

print(head(EuStocksLogReturns))
```

# 2 - Risk-neutralize simulations

## 2 - 1 Yield to maturities (fake *risk-free* rates)

```{r}
ym <- c(0.03013425, 0.03026776, 0.03040053, 0.03053258, 0.03066390, 0.03079450, 0.03092437)

freq <- frequency(EuStocksLogReturns)
(start_preds <- tsp(EuStocksLogReturns)[2] + 1 / freq)
(ym <- stats::ts(ym,
                 start = start_preds,
                 frequency = frequency(EuStocksLogReturns)))
```

## 2 - 2 Risk-neutralized simulations

```{r echo=TRUE, eval=FALSE}
obj <- ahead::ridge2f(EuStocksLogReturns, h = 7L,
                      type_pi = 'bootstrap',
                      B = 10L, ym = ym)
```

```{r echo=TRUE, eval=FALSE}
rowMeans(obj$neutralized_sims$CAC)
```

```{r echo=TRUE, eval=FALSE}
print(ym)
```

```{r echo=TRUE, eval=FALSE}
rowMeans(obj$neutralized_sims$DAX)
```

```{r echo=TRUE, eval=FALSE}
print(ym)
```

# 3 - Visualization

```{r echo=TRUE, fig.width=7.2, eval=FALSE}

#par(mfrow = c(2, 2))

matplot(EuStocksLogReturns, type = 'l', 
     main = "Historical log-Returns", xlab = "time")

plot(ym, main = "fake spot curve", 
     xlab = "time to maturity",
     ylab = "yield", 
     ylim = c(0.02, 0.04))

matplot(obj$neutralized_sims$DAX, type = 'l', 
     main = "simulations of \n predicted DAX log-returns ('risk-neutral')", 
     ylim = c(0.02, 0.04), 
     ylab = "log-returns")

ci <- apply(obj$neutralized_sims$DAX, 1, function(x) t.test(x)$conf.int)
plot(rowMeans(obj$neutralized_sims$DAX), type = 'l', main = "average predicted \n DAX log-returns ('risk-neutral')", col = "blue", 
     ylim = c(0.02, 0.04), 
     ylab = "log-returns")
lines(ci[1, ], col = "red")
lines(ci[2, ], col = "red")
```





