---
title: "Plotting functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`ahead` is a package for **univariate** and **multivariate time series forecasting**, with uncertainty quantification (R and [Python](https://techtonique.github.io/ahead_python/)).

The model used in this demo is `stats::ridge2f`. 

Please remember that _in real life_, this model's hyperparameters will have to be tuned.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Install `ahead`

Here's how to install the R version of the package:

-   **1st method**: from [R-universe](https://techtonique.r-universe.dev)

    In R console:

    ``` r
    options(repos = c(
        techtonique = 'https://techtonique.r-universe.dev',
        CRAN = 'https://cloud.r-project.org'))

    install.packages("ahead")
    ```

-   **2nd method**: from [Github](https://github.com/Techtonique/ahead)

    In R console:

    ``` r
    devtools::install_github("Techtonique/ahead")
    ```

    Or

    ``` r
    remotes::install_github("Techtonique/ahead")
    ```

And here are the packages that will be used in this vignette:

``` r
library(ahead)
library(fpp)
```

```{r include=FALSE}
library(ahead)
library(fpp)
```

# Forecasting and plot predictions

```{r , eval=FALSE}
obj <- ahead::ridge2f(fpp::insurance, h = 7, type_pi = "blockbootstrap", B = 10, 
                      block_length = 5)
```

```{r fig.width=7.2, eval=FALSE}
plot(obj, selected_series = "Quotes", type = "sims", 
     main = "Predictive simulations \n for Quotes")
plot(obj, selected_series = "Quotes", type = "dist", 
     main = "Predictive simulation \n for Quotes")
```

```{r fig.width=7.2, eval=FALSE}
plot(obj, selected_series = "TV.advert", type = "sims", 
     main = "Predictive simulation \n for TV.advert")
plot(obj, selected_series = "TV.advert", type = "dist", 
     main = "Predictive simulation \n for TV.advert")
```

