---
title: "Generalized Linear Model Theta Forecast with attention Pt.3"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generalized Linear Model Theta Forecast with attention Pt.3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r fig.width=7.5}
library(forecast)
library(ahead)

attention_types <- c("dot_product", "scaled_dot_product", 
  "cosine", "exponential", 
  "gaussian", "linear",
  "value_based", "hybrid", 
  "parametric")

results <- data.frame(matrix(0, ncol=length(attention_types), nrow=8))

colnames(results) <- attention_types

rmse <- function(obj, y_test)
{
  sqrt(mean((obj$mean - y_test)**2))
}

y <- AirPassengers

y_split <- misc::splitts(y, split_prob = 0.9)

y_train <- y_split$training

y_test <- y_split$testing

h <- length(y_test)

for (attention_type in attention_types)
{
   misc::debug_print(attention_type)
  
   (obj <- suppressWarnings(ahead::glmthetaf(y_train, h=h, fit_func=MASS::glm.nb, attention = TRUE, attention_type = attention_type)))
  print(obj)
  results[1, attention_type] <- rmse(obj, y_test)
  plot(obj)
  
  (obj <- suppressWarnings(ahead::glmthetaf(y_train, h=h, fit_func=stats::glm, attention = TRUE, attention_type = attention_type)))
  results[2, attention_type] <- rmse(obj, y_test)
  print(obj)
  plot(obj)
  
  (obj <- suppressWarnings(ahead::glmthetaf(y_train, h=h, fit_func=MASS::rlm, attention = TRUE, attention_type = attention_type)))
  results[3, attention_type] <- rmse(obj, y_test)
  print(obj)
  plot(obj)
  
  (obj <- suppressWarnings(ahead::glmthetaf(y_train, h=h, fit_func=MASS::lqs, attention = TRUE, attention_type = attention_type)))
  results[4, attention_type] <- rmse(obj, y_test)
  print(obj)
  plot(obj) 
  
  
  (obj <- suppressWarnings(ahead::glmthetaf(y_train, h=h, fit_func=MASS::glm.nb, attention = FALSE, attention_type = attention_type)))
  print(obj)
  results[5, attention_type] <- rmse(obj, y_test)
  plot(obj)
  
  (obj <- suppressWarnings(ahead::glmthetaf(y_train, h=h, fit_func=stats::glm, attention = FALSE, attention_type = attention_type)))
  results[6, attention_type] <- rmse(obj, y_test)
  print(obj)
  plot(obj)
  
  (obj <- suppressWarnings(ahead::glmthetaf(y_train, h=h, fit_func=MASS::rlm, attention = FALSE, attention_type = attention_type)))
  results[7, attention_type] <- rmse(obj, y_test)
  print(obj)
  plot(obj)
  
  (obj <- suppressWarnings(ahead::glmthetaf(y_train, h=h, fit_func=MASS::lqs, attention = FALSE, attention_type = attention_type)))
  results[8, attention_type] <- rmse(obj, y_test)
  print(obj)
  plot(obj) 

}

print(results)
```


```{r fig.width=7.5}
library(forecast)
library(ahead)
library(MASS)
library(misc)

attention_types <- c(
  "dot_product", "scaled_dot_product", 
  "cosine", "exponential", 
  "gaussian", "linear",
  "value_based", "hybrid", 
  "parametric"
)

fit_funcs <- list(
  MASS::glm.nb,
  stats::glm,
  MASS::rlm,
  MASS::lqs
)

# 8 rows: 4 with attention, 4 without
results <- matrix(0, nrow = 8, ncol = length(attention_types))
colnames(results) <- attention_types
rownames(results) <- c(
  paste0("att_", c("glm.nb", "glm", "rlm", "lqs")),
  paste0("noatt_", c("glm.nb", "glm", "rlm", "lqs"))
)
results <- as.data.frame(results)

rmse <- function(obj, y_test) {
  sqrt(mean((obj$mean - y_test)^2))
}

y <- AirPassengers
y_split <- misc::splitts(y, split_prob = 0.9)
y_train <- y_split$training
y_test  <- y_split$testing
h <- length(y_test)

for (attention_type in attention_types) {
  misc::debug_print(attention_type)

  for (i in seq_along(fit_funcs)) {
    fit_func <- fit_funcs[[i]]

    # With attention
    obj <- suppressWarnings(
      ahead::glmthetaf(
        y_train, h = h, fit_func = fit_func,
        attention = TRUE, attention_type = attention_type
      )
    )
    misc::debug_print(paste("With attention -", deparse(substitute(fit_func))))
    print(obj)
    results[i, attention_type] <- rmse(obj, y_test)
    plot(obj)

    # Without attention
    obj <- suppressWarnings(
      ahead::glmthetaf(
        y_train, h = h, fit_func = fit_func,
        attention = FALSE, attention_type = attention_type
      )
    )
    misc::debug_print(paste("No attention -", deparse(substitute(fit_func))))
    print(obj)
    results[i + 4, attention_type] <- rmse(obj, y_test)
    plot(obj)
  }
}

print(results)

